Spirals launched by a planet in a shadowed disk

### 2024-04-13 src/tex/ms.tex
finish setup text

\section{Methods}
\label{sec:methods}
\subsection{Simulation setup}
\label{subsec:setup}
We use the magneto-hydrodynamics code Athena++\footnote{\url{https://github.com/PrincetonUniversity/athena}{https://github.com/PrincetonUniversity/athena}} 
\citep{stone2020} version 21.0 \citep{athena++developmentteam2021} to simulate a 2D gas disk. The simulation was set up in cylindrical coordinates with 149
cells in the $\ln{r}$ direction with $r \in [0.4, 2.5]$ and 512 cells in the $\phi \in [-\pi, \pi]$ direction. The inner and outer boundaries have a user-defined fixed boundary
condition and the azimuthal boundary has a periodic boundary condition to ensure material can orbit the central star indefinitly. The dimensionality of the problem
is set up so that all masses are in $M_\odot$, distances are in AU, and times are in year/$2\pi$. This means that a planet placed in a circular orbit at $r=1$
will have an orbital period of $2\pi$.

The disk is set up so that $\rho = 1$ at $r=1$ and has the radial profile $\rho \propto r^{-1}$. The gas in the disk initially moves at the local keplerian velocity.
A shadow on the disk is prescribed by setting the sound speed -- analogous to setting the temperature as 
\begin{equation}
    c_s = \frac{\partial P}{\partial \rho} \propto T
\end{equation}
We ensure that the shadow does not move with the gas in its orbit by setting the cooling timescale to $10^{-6} \ll 1$. The sounds speed is essentially reset each time
the grid is updated.

The shadow is parameterized by three numbers: \texttt{h\_reduced\_factor}, \texttt{half\_shadow}, and \texttt{half\_unshadow},
which we will write here as $h$, $a$, and $b$, respectively. The former is the dimensionless magnitude of the sound speed reduction
and the latter two are the half-width of the regions with constant sound speed (i.e. no the transition region) and have
units of radians.

@@

\begin{equation*}
    \label{eq:H}
    \begin{array}{cccc}
        H = & h & \text{if} & (|\phi|~\%~\pi) \le a  \\
            & 1 & \text{if} & (\left | |\phi| - \frac{\pi}{2}\right |~\%~\pi) \le b \\
            & h + (1-h) \sin{\left ( \frac{\pi}{2}  \frac{|\phi| - b}{\frac{\pi}{2} - a - b} \right )} & & \text{otherwise}
    \end{array}
\end{equation*}

\begin{figure}
    % \includegraphics[width=0.5\textwidth]{}
    \caption{Initial gas temperature maps for our three setups. {\bf Left:} In the case where there is no shadow the temperature is azimuthally symmetric.
    {\bf Center:} In our ``narrow'' case we see a deep shadow covers a narrow region of the disk. This is analogous to a thin, optically thick inner disk.
    {\bf Right:} In our ``wide'' case we see a much shallower shadow that covers a significant portion of the disk. In this case $a,b=0$ and the third line of
    Equation \ref{eq:H} becomes a sine function for all $\phi$.
    }
\end{figure}

\begin{table}
    \begin{tabular}{cccc}
        \hline
        \hline
        Name & Width & $c_{sh}/c_{un}$ & $m_p$ \\
        \texttt{no\_shadow} & None & 0 & $10^{-4}$ \\

        \hline

    \end{tabular}
\end{table}

>>>>

\section{Methods}
\label{sec:methods}
\subsection{Simulation setup}
\label{subsec:setup}
We use the magneto-hydrodynamics code Athena++\footnote{\url{https://github.com/PrincetonUniversity/athena}{https://github.com/PrincetonUniversity/athena}} 
\citep{stone2020} version 21.0 \citep{athena++developmentteam2021} to simulate a 2D gas disk. The simulation was set up in cylindrical coordinates with 149
cells in the $\ln{r}$ direction with $r \in [0.4, 2.5]$ and 512 cells in the $\phi \in [-\pi, \pi]$ direction. The inner and outer boundaries have a user-defined fixed boundary
condition and the azimuthal boundary has a periodic boundary condition to ensure material can orbit the central star indefinitly. The dimensionality of the problem
is set up so that all masses are in $M_\odot$, distances are in AU, and times are in year/$2\pi$. This means that a planet placed in a circular orbit at $r=1$
will have an orbital period of $2\pi$. We chose a planet mass of $10^{-4}$ in order to create a detectable effect in the presence of noise from the shadow and
to not stray too far from the linear spiral regime.

The disk is set up so that $\rho = 1$ at $r=1$ and has the radial profile $\rho \propto r^{-1}$. The gas in the disk initially moves at the local keplerian velocity.
A shadow on the disk is prescribed by setting the sound speed -- analogous to setting the temperature as 
\begin{equation}
    c_s = \frac{\partial P}{\partial \rho} \propto T
\end{equation}
We ensure that the shadow does not move with the gas in its orbit by setting the cooling timescale to $10^{-6} \ll 1$. Essentially, the temperature is reset each time
the grid is updated.

The shadow is parameterized by three numbers: \texttt{h\_reduced\_factor}, \texttt{half\_shadow}, and \texttt{half\_unshadow},
which we will write here as $h$, $a$, and $b$, respectively. The former is the dimensionless magnitude of the sound speed reduction
and the latter two are the half-width of the regions with constant sound speed (i.e. not the transition region) and have
units of radians.

@@

\begin{equation*}
    \label{eq:H}
    \begin{array}{cccc}
        H = & h & \text{if} & |\sin{\phi}| \le \sin{a}  \\
            & 1 & \text{if} & |\cos{\phi}| \le \sin{b} \\
            & h + \frac{1-h}{2} \left (  1-\cos{\left ( \frac{2 \pi | \phi + m\pi |}{\pi - 2a - 2b}  \right )  } \right ) & & \text{otherwise}
    \end{array}
\end{equation*}
and
\begin{equation}
    \begin{array}{cccc}
        m = & 1 & \text{if} & \phi < -\frac{\pi}{2} \\
            & -1 & \text{if} & \phi > \frac{\pi}{2} \\
            & 0 & & \text{otherwise}
    \end{array}
\end{equation}
Figure \ref{fig:cs} shows the sound speed profiles used in this work.

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5in]{figures/sound_speed.pdf}
        \caption{Sound speed as a function of azimuthal angle. Here we show how the sound speed is prescribed in our two science cases. The cases
        share a minimum and maximum sound speed, but the transitions between these values are very different.}    
    \end{center}
    \label{fig:cs}
\end{figure}

\begin{figure}
    % \includegraphics[width=0.5\textwidth]{}
    \caption{Initial gas temperature maps for our three setups. {\bf Left:} In the case where there is no shadow the temperature is azimuthally symmetric.
    {\bf Center:} In our ``narrow'' case we see a deep shadow covers a narrow region of the disk. This is analogous to a thin, optically thick inner disk.
    {\bf Right:} In our ``wide'' case we see a much shallower shadow that covers a significant portion of the disk. In this case $a,b=0$ and the third line of
    Equation \ref{eq:H} becomes a sine function for all $\phi$.
    }
    \label{fig:setup}
\end{figure}

We will investigate two scenarios:
\begin{enumerate}
    \item A narrow shadow, analogous to that produced by a thin inner disk with a scale height of $\frac{h}{r} = 0.05$. We chose a value of $a$ based on the assumption that the inner disk
        is optically thick up to 3 scale heights. We call this case ``narrow''.
    \item A wide shadow, like that produced by a diffuse inner disk with a very high scale height. In this limit the temperature is simply a sinusoidal function of $\phi$. We call this case ``wide''.
\end{enumerate}
In order to adequately investigate these scenarios we must run a total of five simulations, described in Table \ref{tab:scenarios}. For each of our two science cases we run an
additional simulation with no planet in order to subtract out effects that are purely due to the shadow. Finally, we also run a control simulation without any shadow.
Figure \ref{fig:setup} shows the initial condition of each case described here.

\begin{table}
    \label{tab:scenarios}
    \begin{tabular}{ccccc}
        \hline
        \hline
        Name & $a$ & $b$ & $h$ & $m_p$ \\
        \texttt{no\_shadow} & - & - & 0 & $10^{-5}$ \\
        \texttt{narrow\_with} & 0.15 & 1.32 & 0.9 & $10^{-5}$ \\
        \texttt{narrow\_without} & 0.15 & 1.32 & 0.9 & 0 \\
        \texttt{wide\_with} & 0 & 0 & 0.9 & $10^{-5}$ \\
        \texttt{wide\_with} & 0 & 0 & 0.9 & 0 \\
        \hline
    \end{tabular}
    \caption{Simulation parameters. The planet mass of $10^{-5}$ is analogous to a planet three times as massive as Earth.}
\end{table}
